Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-04-09 docs/ACM/paper.tex
Adding discussion on versatility.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_Versatility}
\caption{{\Taskminer}'s performance in the real world. This chart shows that our algorithm also performs
well on programs that are not deemed inherently parallel, being able to find hidden parallelism
in 20 programs from LLVM's Test Suite.}
\label{fig:TM_Versatility}
\end{center}
\end{figure}

Our tool has also performed well on \textsf{LLVM's Test Suite}, being able to even find hidden parallelism.
From a total of XX single source programs,
{\Taskminer} automatically produced parallel annotated versions of 20 of them, as seen in Figure
\ref{fig:testsuite}. We were able to not only find hidden parallelism in those programs, which are not
considered inherently parallel, but also produce these parallel versions automatically. Amongst
those programs, one caught our attention: \textit{misc/lowercase.c}, which applies a function to every
string in the text. The automatically annotated version of this program was 3.5 times faster than
the sequential one. We have also nice speed up numbers in \textit{misc/ffbench} and \textit{misc/mandel-2}.

\begin{table}[h]
\centering
\caption{Different kinds of tasks among the BOTS benchmarks. {\Taskminer}'s versaitily allows the user to define which type of tasks should be mined, annotated or cared for.}
\label{tab:types}
\begin{tabular}{|r|c|c|c|}
\hline
                   & \textbf{Recursive} & \textbf{FunctionCall} & \textbf{Region} \\ \hline
\textbf{fft}       & \checkmark         &                       & \checkmark      \\ \hline
\textbf{fib}       & \checkmark         &                       &                 \\ \hline
\textbf{floorplan} & \checkmark         &                       & \checkmark      \\ \hline
\textbf{sort}      & \checkmark         &                       & \checkmark      \\ \hline
\textbf{sparselu}  &                    & \checkmark            & \checkmark      \\ \hline
\textbf{strassen}  & \checkmark         &                       & \checkmark      \\ \hline
\textbf{uts}       & \checkmark         &                       &                 \\ \hline
\textbf{nqueens}   & \checkmark         &                       & \checkmark      \\ \hline
\textbf{knapsack}  & \checkmark         &                       & \checkmark      \\ \hline
\textbf{health}    & \checkmark         & \checkmark            & \checkmark      \\ \hline
\textbf{jacobi}    &                    &  \checkmark                     &                 \\ \hline
\end{tabular}
\end{table}

{\Taskminer}'s versatility also arises in regards to the diversity of types of task parallelism that it
can find in a program. Table \ref{tab:types} shows all types of tasks found in our main benchmarks.
There are 3 types of tasks: recursive, which arises from vanes in Divide and Conquer
algorithms; function call, which arises within loops that are not clearly parallel; and region, which 
are similar to function calls, but instead are entire fragments of code within loops that can be executed in
parallel. We provide the user complete decision power to mine for any type of tasks they wish when
running our tool.

>>>>

The benchmarks used in Section~\ref{sub:performance} have been written to
demonstrate the power of different parallel systems; hence, they have already
been written in a way that simplifies the discovery of parallelism in programs.
In this section, we address the following question: ``can \Taskminer" find
parallelism in general programs?
To answer it, we have applied \Taskminer{} onto the 219 C programs available in
the LLVM test suite, and have compared the annotated version
against the sequential version compiled with \textsf{clang} -O3.
\Taskminer{} annotates a benchmark if it can find symbolic bounds to every
memory access (load or store) used within a region marked as a task candidate.
Under this constraint, we have been able to discover tasks in 63 benchmarks.
Figure~\ref{fig:TM_Versatility} relates the number of tasks and the number of
instructions in benchmarks.
To avoid counting multiple C files for the same benchmark,
Figure~\ref{fig:TM_Versatility} contains only benchmarks that are
constituted by a single file (present in LLVM's \textsf{SingleSource} folder).
As this experiment demonstrates, \Taskminer{} {\em can annotate a non-trivial
number of real-world benchmarks}.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_Versatility}
\caption{Relation between number of task regions and program size.
Each benchmark is a complete C file.}
\label{fig:TM_Versatility}
\end{center}
\end{figure}

Most of the benchmarks, e.g., 27, that have been automatically annotated did not
give us speedups, because the regions marked as tasks were too small to influence
the program's runtime.
We have observed slowdowns in 17 benchmarks.
In this case, interactions between data-structures end up forcing the OpenMP
runtime to serialize execution of tasks.
And we have measured speedups above 5\% in 19 benchmarks.
In one case, \textsf{Misc/lowercase}, the speedup is above 3.5x, and in
three cases, it goes beyond 1.5x.
Figure~\ref{fig:TM_GeneralProgs} summarizes these findings.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_GeneralProgs}
\caption{20 best speedups obtained by \Taskminer{} when applied onto the LLVM
test suite.
The larger the bar, the better.
\textbf{LoC} stands for ``Lines of C".}
\label{fig:TM_GeneralProgs}
\end{center}
\end{figure}

%{\Taskminer}'s versatility also arises in regards to the diversity of types of task parallelism that it
%can find in a program. Table \ref{tab:types} shows all types of tasks found in our main benchmarks.
%There are 3 types of tasks: recursive, which arises from vanes in Divide and Conquer
%algorithms; function call, which arises within loops that are not clearly parallel; and region, which 
%are similar to function calls, but instead are entire fragments of code within loops that can be executed in
%parallel. We provide the user complete decision power to mine for any type of tasks they wish when
%running our tool.
